\chapter{Related Work}

\section{Ixy}
Ixy is a network interface card (NIC) driver for Intel's 82599 10GbE NICs (ixgbe family) \cite{emmerich2019user}.
Ixy has been implemented in many languages, e.g. C, Go and Rust. Ixy.rs is the Rust implementation of the Ixy driver \cite{baellmann}.
Stefan Huber implemented IOMMU support for the Ixy.rs driver. It was concluded that while using the IOMMU with \qty{2}{\mebi\byte} pages, the performance matches the performance without the IOMMU. On the other hand, using \qty{4}{\kibi\byte} pages leads to a potential 75\% performance loss. Additionally it was found that the tested Intel Xeon E5-2620 v3 6-core CPU IOMMU TLB has a maximum size of 64 entries \cite{iommuhuber}. Rolf Neugebauer et al. determined the same 64 IOTLB size on the tested Intel Xeon E5-2630v4 2.2GHz \cite{pcieperfnegebauer}.

\section{Data Plane Development Kit}
The Data Plane Development Kit (DPDK) is a framework for developing user-space network card drivers. It allows for high performance network applications. It can either run using direct memory access with physical addresses or using VFIO \cite{aboutdpdk}. DPDK offers polling drivers for a variety of network cards.
It is one of the most successful projects in the world of userspace drivers and has influenced many advances in the IOMMU space.

\section{Storage Performance Development Kit}
The demand for high-speed userspace drivers in storage applications inspired the development of the Storage Performance Development Kit (SPDK). SPDK uses some shared libraries and architecture with DPDK. Especially through the wide adoption of the NVMe protocol and the standardization of said protocol, only one driver for all NVMe SSDs has to be developed.
NVMe is a storage protocol which is widely used, modern and highly performant. Therefore it is a protocol for which many drivers have been written, including userspace drivers. The Storage Performance Development Kit (SPDK) provides ``a collection of tools and libraries for writing high performance, scalable, user-mode storage applications'' \cite{spdkindex}. It includes an user-space NVMe driver which is fast and production-ready. While this driver supports the use of the driver without the IOMMU, the SPDK Documentation recommends using the IOMMU as using VFIO and the IOMMU is the "future proof...long-term foundation" for SPDK \cite{spdkmemory}. Even though SPDK is the established userspace NVMe driver option, the drawbacks include its high complexity even for simple applications, as well as it being written in C.
